SIR Model With Testing Flows and Isolation

### 2020-11-06 codes/ali_test_sir.Rmd
update analysis of R0 wrt omega

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, 
             W_S=1, W_I=1, W_R=1, #weights
             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2 #eta_w/eta_t
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
# rootfun <- function (time, state,params) { 
#   unpack(as.list(c(state,params)))
#   return(I_u - 1) }
```

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

4. Taylor expansion gives $R_0|_{(\rho=0,W_S=1)} = \beta/\gamma$.

5. Taylor expansion gives $R_0|_{(\rho=0,W_S=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_I+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma+\omega p_I)W_I+ \gamma \rho(\gamma+\omega)}$.

@@

>>>>

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, #0.8, 1/3
             W_S=1, W_I=1, W_R=1, #weights
             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2 #eta_w/eta_t
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
# rootfun <- function (time, state,params) { 
#   unpack(as.list(c(state,params)))
#   return(I_u - 1) }
```

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

In general, $a \geq 0$ is not true. $$a \geq 0 \iff (s-p_I) \eta_t \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_t} \geq 1$. As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.

**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_t=0$. This leads to $a \leq 0$, $b<0$, $c<0$, thus $\partial{R_0}/\partial{\omega} \geq 0$. Intuitively, returning the test results back to the awaiting individuals while they are in a perfect isolation results in an increase of $R_0$.

@@

```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
R0_omega <- function(params){
  unpack(as.list(params))
         # a <- (eta_w-1)*W_I-(eta_t-1)*p_I*W_I
         a <- ((s-p_I)*eta_t + (p_I-1))*W_I  #involving s
         b <- 2*(eta_w-1)*gamma*W_S
         c <- b*gamma/2

         d <- (-beta*rho)/(gamma*W_S*omega^2*(gamma+omega)^2)*(a*omega^2 + b*omega + c)
         return(d)
}

R0_omega(params)

```{r }
 # specify the ranges
if(F){
beta  <- 0.5  ## set beta 
gamma <- seq(0.1,0.5, length.out=n_out)
p_I <- 0.9

# eta_w <- seq(0,1, length.out=n_out)
eta_t <- seq(0.001,1, length.out=n_out)
s <- seq(1,2,length.out=n_out)

rho <- seq(0,0.01, length.out=n_in)
omega <- seq(0.011,2 , length.out=n_in) #note omega must be > rho, was 0.011

W_S <- 1/100
W_I <- 1/3
W_R <- 1/3

 df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
                    W_S=W_S,W_I=W_I,W_R=W_R,
                    p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

df3_2 <- data.frame(df3_1,R0=apply(df3_1,1,function(params_in)R02(params=params_in)),
                      R0_omega= apply(df3_1,1,function(params_in)R0_omega(params=params_in))
                       )
}
```

```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
if(F){
p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p3
    + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=0,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="dR[0]"))
)
}
```
